Uncovering underlying high-level musical content in the time domain \\ \small Leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn deep audio embeddings with applications to boundary detection tasks

### 2023-07-06 acknowledgements/acknowledgements.tex
Updates from Overleaf

\begin{itemize}
\item Carl Thomé and Carlos Lordelo, whose expertise in MIR, ML, and DSP has been pivotal to the progress and success of my thesis. Their invaluable guidance and unwavering encouragement have empowered me to overcome the challenges encountered throughout this journey. Their dedication and commitment to mentorship are truly inspiring, and their influence has left an indelible mark on my academic path. The opportunity to work under their mentorship has been a privilege and an honor. Their profound impact is beyond what words can encapsulate.

\vspace*{3mm}
\item Cody Hesse and Sebastian Löf

\vspace*{3mm}
\item Oriol Nieto and J. Smith

\vspace*{3mm}
\item Epidemic Sound

\vspace*{3mm}
\item The Music Technology Group (MTG) and Universitat Pompeu Fabra (UPF). My profound thanks go to Xavier Serra for his faith in accepting a non-traditional profile to the master's program. Additionally, I appreciate Pedro Ramoneda and Genís Plaja for their patient assistance throughout my journey.

\vspace*{3mm}
\item My colleague at My Sheet Music Transcriptions, Oriol López, for his exceptional support and flexibility throughout my thesis. His understanding effectively allowed me to balance my work and research responsibilities.

@@

Thank you all for playing a crucial role in completing my thesis. I sincerely appreciate everything you have done for me and will always cherish the knowledge and experience I have gained under your guidance.

>>>>

\begin{itemize}
\item Carl Thomé and Carlos Lordelo, whose expertise in MIR, ML, and DSP has been pivotal to the success of my thesis. Their guidance and encouragement have enabled me to overcome the challenges encountered during this journey. Their commitment to mentorship is truly inspiring, and their influence has significantly shaped my academic path. 
Working under their mentorship has been both a privilege and an honor.

\vspace*{3mm}
\item Cody Hesse and Sebastian Löf. Their generosity in sharing their time, knowledge, and insights has critically enhanced the quality of this research work.

\vspace*{3mm}
\item The Music Technology Group (MTG) and Universitat Pompeu Fabra (UPF). My profound thanks go to Xavier Serra for his faith in accepting a non-traditional profile into the master's program. I am equally grateful to Pedro Ramoneda and Genís Plaja for their consistent support throughout my journey.

@@

I sincerely thank you all for your integral roles in completing my thesis. I deeply value the knowledge and experience gained under your guidance and will cherish it always.

### 2023-07-06 discussion/discussion.tex
Updates from Overleaf

While our musically-informed technique does not yet outperform the existing state-of-the-art baselines \cite{deepfeaturesegment, SalamonDeepSegmentation}, it exhibits significant potential in boundary detection tasks, particularly when expanding the training set. The improvement we've observed between datasets is noteworthy, and when compared with two of the acoustic features, our results are highly competitive. We've also managed to circumvent the typical issues associated with dataset enlargements, such as the need for extensive supervision or human annotation, which gives our method an edge in terms of practicality and scalability effectively turning what is usually seen as an expensive hurdle into a much more manageable task.

@@

Too little representation layer?

Too little model for raw audio input? \cite{verydeep}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/images/Mean group aggregation.png}
    \caption{Caption}
    \label{fig:enter-label}
\end{figure}

>>>>

While our musically-informed technique does not yet outperform the existing state-of-the-art baselines \cite{deepfeaturesegment, SalamonDeepSegmentation}, it exhibits significant potential in boundary detection tasks, particularly when expanding the training set. The improvement we've observed between datasets is noteworthy, and when compared with two of the acoustic features, our results are highly competitive. We've also managed to circumvent the typical issues associated with dataset enlargements, such as the need for extensive supervision or human annotation, which gives our method an edge in terms of practicality and scalability, effectively turning what is usually seen as an expensive hurdle into a much more manageable task.

@@

We argue whether factors such as the size of the representation layer and the size of the raw audio model might be insufficient \cite{verydeep}. Furthermore, we posit that with additional time to iterate repeatedly over the dataset, the loss function could decrease further with minimal risk of overfitting. \ref{fig:lossf}

While our results may not yet rival state-of-the-art benchmarks, they offer an encouraging starting point. They affirm we're on the right path and reveal the potential of this research direction. Importantly, we've seen that progress doesn't have to be expensive—though time and potential copyright issues are considerations.

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{figures/images/Mean group aggregation.png}
    \caption[Loss function performance plot]{Mean group aggregation of the decrease of the loss function over trainer/global steps. Training loss against validation loss.}
    \label{fig:lossf}
\end{figure}

### 2023-07-06 introduction/Subsections/Assumptions.tex
Updates from Overleaf

This thesis posits that enduring, high-level musical elements persist regardless of sonic changes and constantly evolve within their context, mirroring the nature of sheet music's symbolic domain. A single waveform can correspond to a finite set of representations. Still, one music piece -or music representation- can yield infinite interpretations, varying by factors such as instruments and performance styles, to name just a few.

>>>>

This thesis posits that enduring, high-level musical elements persist regardless of sonic changes and constantly evolve within their context, mirroring the nature of sheet music's symbolic domain. A single waveform can correspond to a finite set of representations. Still, one music piece ---or music representation--- can yield infinite interpretations, varying by factors such as instruments and performance styles, to name just a few.

### 2023-07-06 methods/subsections/implementation.tex
Updates from Overleaf

\subsubsection{Encoder architecture: SampleCNN}

@@

Circling back to our premises and assumptions, we posit that high-level musical content unfolds over time; therefore, we argue that the temporal structure of the negative images, in contrast to our anchor, should be disrupted. While maintaining similar sonic qualities, the content should be rendered unintelligible.

>>>>

\subsubsection{Encoder: SampleCNN}

@@

Circling back to our premises and assumptions, we posit that high-level musical content unfolds over time; therefore, we argue that the temporal structure of the negative images, in contrast to our anchor, must be disrupted. While maintaining similar sonic qualities, the content should be rendered unintelligible.

### 2023-07-06 results/subsections/tables.tex
Updates from Overleaf

\begin{table}[h]
\centering
\small
\begin{tabularx}{\textwidth}{>{\centering\arraybackslash}p{4.5cm}>{\centering\arraybackslash}X>{\centering\arraybackslash}X>{\centering\arraybackslash}X}
\toprule
\textbf{Training dataset [Ref]}  & \textbf{Precision} & \textbf{Recall} & \textbf{F-measure} \\
\midrule
\addlinespace
GTZAN \cite{GTZAN} & 0.228 & 0.171 & 0.185 \\
\addlinespace
MSD \cite{MSD} &  \textbf{0.333} &  \textbf{0.280} & \textbf{0.288} \\
\addlinespace
\bottomrule
\end{tabularx}
\caption[Caption]{\small{Comparison of precision, recall, and f-measure for GTZAN-trained versus MSD-trained embeddiograms on SALAMI dataset computed using the Structural Feature (SF)\cite{sf} algorithm.}}
\label{tab:comparison_table}
\end{table}

@@

\begin{table}[ht]
\begin{threeparttable}
\centering
\small
\begin{tabularx}{\textwidth}{
  >{\centering\arraybackslash}p{4.5cm}
  >{\centering\arraybackslash}X
  >{\centering\arraybackslash}X
  >{\centering\arraybackslash}X}
\toprule
\thead{\textbf{Authors [Ref], Year}} & \thead{\textbf{Input}\tnote{1}} & \thead{\textbf{Method}} & \thead{\textbf{F-measure}} \\
\midrule
Turnbull et al. \cite{Turnbull2007ABOOSTING}, 2007 & MFCCs, chromas, spectrogram & Boosted Decision Stump  & 0.378 \\
\addlinespace
Kaiser et al. \cite{27}, 2012 & SSM & Novelty measure  & 0.286 \\
\addlinespace
McFee \& Ellis \cite{20}, 2013 & MLS & Fisher’s Linear Discriminant  & 0.317 \\
\addlinespace
Nieto \& Bello \cite{28}, 2014 & MFCCs, chromas & Checkerboard-like kernel  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2015 & Timbre-type histograms & HMM  & 0.213 \\
\addlinespace
Nieto \cite{30}, 2016 & CQT Spectrogram & Linear Discriminant Analysis  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2017 & Timbre-type histograms & HMM  & 0.212 \\
\addlinespace
Sargent et al. \cite{34}, 2011 & MFCCs, chromas & Viterbi  & 0.356 \\
\addlinespace
Ullrich et. al \cite{22}, 2014 & MLS & CNN  & 0.465 \\
\addlinespace
Grill \& Schlüter \cite{4}, 2015 & MLS, SSLMs & CNN  & 0.523 \\
\addlinespace
Grill \& Schlüter \cite{Grill2015MusicAnnotations}, 2015 & MLS, PCPs, SSLMs & CNN  & 0.508 \\
\addlinespace
Hadria \& Peeterss \cite{35}, 2017 & MLS, SSLMs & CNN  & 0.291 \\
\addlinespace
McCallum \cite{deepfeaturesegment}, 2019 & CQT & CNN  & \textbf{0.535} \\
\addlinespace
Ours, 2023 & Raw waveforms & CNN  & 0.288 \\
\bottomrule
\end{tabularx}
\caption[Baseline. State-of-the-art table.]{Previous studies' boundary detection f-measure results using unsupervised methods for a 0.5s time-window tolerance. Only the top-performing algorithm for each year on the SALAMI dataset is displayed. This table has been extended from \cite{Hernandez-Olivan2021MusicFeatures}.}
\label{tab:comparison_table}
\begin{tablenotes}\footnotesize
\item[1] \textbf{Legend:} SSM: Self-Similarity Matrix, MLS: Mel Spectrogram, MFCC: Mel-Frequency Cepstral Coefficient, CQT: Constant Q Transform, PCP: Pitch Class Profile, SSLM: Self-Similarity Lag Matrix.
\end{tablenotes}
\end{threeparttable}
\end{table}

>>>>

\begin{table}[h]
\centering
\small
\begin{tabularx}{\textwidth}{>{\centering\arraybackslash}p{4.5cm}>{\centering\arraybackslash}X>{\centering\arraybackslash}X>{\centering\arraybackslash}X}
\toprule
\textbf{Training dataset [Ref]}  & \textbf{Precision} & \textbf{Recall} & \textbf{F-measure} \\
\midrule
\addlinespace
GTZAN \cite{GTZAN} & 0.228 & 0.171 & 0.185 \\
\addlinespace
MSD \cite{MSD} &  \textbf{0.333} &  \textbf{0.280} & \textbf{0.288} \\
\addlinespace
\bottomrule
\end{tabularx}
\caption[GTZAN-trained versus MSD-trained embeddiograms]{\small{Comparison of precision, recall, and f-measure for GTZAN-trained versus MSD-trained embeddiograms on SALAMI dataset computed using the Structural Feature (SF)\cite{sf} algorithm.}}
\label{tab:comparison_table}
\end{table}

@@

\begin{table}[ht]
\begin{threeparttable}
\centering
\small
\begin{tabularx}{\textwidth}{
  >{\centering\arraybackslash}p{4.5cm}
  >{\centering\arraybackslash}X
  >{\centering\arraybackslash}X
  >{\centering\arraybackslash}X}
\toprule
\thead{\textbf{Authors [Ref], Year}} & \thead{\textbf{Input}\tnote{1}} & \thead{\textbf{Method}} & \thead{\textbf{F-measure}} \\
\midrule
Turnbull et al. \cite{Turnbull2007ABOOSTING}, 2007 & MFCCs, chromas, spectrogram & Boosted Decision Stump  & 0.378 \\
\addlinespace
Kaiser et al. \cite{27}, 2012 & SSM & Novelty measure  & 0.286 \\
\addlinespace
Sargent et al. \cite{34}, 2011 & MFCCs, chromas & Viterbi  & 0.356 \\
\addlinespace
McFee \& Ellis \cite{20}, 2013 & MLS & Fisher’s Linear Discriminant  & 0.317 \\
\addlinespace
Nieto \& Bello \cite{28}, 2014 & MFCCs, chromas & Checkerboard-like kernel  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2015 & Timbre-type histograms & HMM  & 0.213 \\
\addlinespace
Nieto \cite{30}, 2016 & CQT Spectrogram & Linear Discriminant Analysis  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2017 & Timbre-type histograms & HMM  & 0.212 \\
\addlinespace
Ullrich et. al \cite{22}, 2014 & MLS & CNN  & 0.465 \\
\addlinespace
Grill \& Schlüter \cite{4}, 2015 & MLS, SSLMs & CNN  & 0.523 \\
\addlinespace
Grill \& Schlüter \cite{Grill2015MusicAnnotations}, 2015 & MLS, PCPs, SSLMs & CNN  & 0.508 \\
\addlinespace
Hadria \& Peeterss \cite{35}, 2017 & MLS, SSLMs & CNN  & 0.291 \\
\addlinespace
McCallum \cite{deepfeaturesegment}, 2019 & CQT & CNN  & \textbf{0.535} \\
\addlinespace
Ours, 2023 & Raw waveforms & CNN  & 0.288 \\
\bottomrule
\end{tabularx}
\caption[Baseline. State-of-the-art table.]{Previous studies' boundary detection f-measure results using unsupervised methods for a 0.5s time-window tolerance. Only the top-performing algorithm for each year on the SALAMI dataset is displayed. This table has been extended from \cite{Hernandez-Olivan2021MusicFeatures}.}
\label{tab:comparison_table}
\begin{tablenotes}\footnotesize
\item[1] \textbf{Legend:} SSM: Self-Similarity Matrix, MLS: Mel Spectrogram, MFCC: Mel-Frequency Cepstral Coefficient, CQT: Constant Q Transform, PCP: Pitch Class Profile, SSLM: Self-Similarity Lag Matrix.
\end{tablenotes}
\end{threeparttable}
\end{table}
